---
title: "R Notebook"
output: html_notebook
---

```{r}
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr)
library(tidyverse)
library(ggplot2)
```

```{r}
#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list
```



```{r}
one_play <- function(state){
    #browser() ## what's this?
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}
```


```{r}
#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE


stop_play <- function(state){
  #browser()
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}
```


```{r}
#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 

```


```{r}
# Simulation
walk_out_money <- rep(NA, 100)
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_last
}


# Walk out money distribution
hist(walk_out_money, breaks = 10)

# Estimated probability of walking out with extra cash
mean(walk_out_money > 200)

# Estimated earnings
mean(walk_out_money - 200)
```

### 1. In Figure 1, I demonstrate some examples of grambler earnings over a series of wagers at the roulette wheel.

```{r}
# Q1
set.seed(2)
earning1 <- one_series(B = 200, W = 300, L = 1000, M = 100) - 200
earning2 <- one_series(B = 200, W = 300, L = 1000, M = 100) - 200
earning3 <- one_series(B = 200, W = 300, L = 1000, M = 100) - 200
plot(earning1, main = "Figure 1: Gamblers earnings / losses over time", xlab = "play number", 
     ylab = "earnings", type ="b", col = "red")
lines(earning2, xlab = "play number", ylab = "earnings", col = "blue")
lines(earning3, xlab = "play number", ylab = "earnings", col = "green")
```



```{r}
# in class code

set.seed(2)
s1 <- one_series(B = 200, W = 300, L = 1000, M = 100) - 200
plot(s1[1:120], type = "b")

set.seed(3)
s1 <- one_series(B = 200, W = 300, L = 1000, M = 100) - 200
plot(s1)

```

### 2.1. When we set the budget from 10 to 300 (budget has to be lower than winning threshold), average earnings will first decrease and then increase with the increase of budget. 

```{r}
# change the budget
walk_out_money <- rep(NA, 100)
b <- matrix(nrow = 100, ncol = 30)


# For loop eg:    for (i in 1:100) {...}

for(i in seq(1, 30)){
  for(j in seq_along(walk_out_money)){
    b[j, i] <- (one_series(B= 10*i, W = 300, L = 1000, M = 200) - 10*i) %>% get_last
  }
}

# Estimated earnings
mean_earnings <- colMeans(b)

df1 <- as.data.frame(mean_earnings)
df1["Budget"] <- seq(10,300,10)
df1 %>%
  ggplot(ggtitle("Figure 2: Change the budget"), mapping = aes(x = Budget, y = mean_earnings)) +
  geom_point() +
  geom_smooth()

# plot earnings
#plot(x = seq(10, 300, 10), y = mean_earnings, xlab = "budget", type = "l")

```


### 2.2 When we set the winning threshold from 220 to 840, the average earnings will decrease.

```{r}
# change the winning

walk_out_money <- rep(NA, 100)
w <- matrix(nrow = 100, ncol = 32)

for(i in seq(1, 32)){
  for(j in seq_along(walk_out_money)){
    w[j, i] <- one_series(B= 200, W = 200 + 20 * i, L = 1000, M = 200) %>% get_last
  }
}

# Estimated earnings
mean_earnings <- colMeans(w) - 200

df2 <- as.data.frame(mean_earnings)
df2["Winning_threshold"] <- seq(220,840,20)
df2 %>%
  ggplot(mapping = aes(x = Winning_threshold, y = mean_earnings)) +
  geom_point() +
  geom_smooth()

# plot earnings
#plot(mean_earnings)
```


### 2.3 When we set the Time threshold from 0 to 620, we can see that there is a sharp decrease before 300 plays. After 300 plays, the average earnings are stable.

```{r}
# change time threshold
walk_out_money <- rep(NA, 100)
l <- matrix(nrow = 100, ncol = 32)

for(i in seq(1, 32)){
  for(j in seq_along(walk_out_money)){
    l[j, i] <- one_series(B= 200, W = 300, L = 20 * (i-1), M = 200) %>% get_last
  }
}

# Estimated earnings
mean_earnings <- colMeans(l) - 200

df3 <- as.data.frame(mean_earnings)
df3["Time_threshold"] <- seq(0,620,20)
df3 %>%
  ggplot(mapping = aes(x = Time_threshold, y = mean_earnings)) +
  geom_point() +
  geom_smooth()

# plot earnings
#plot(mean_earnings)
```

### 2.4 When we set Maximum wagers from 0 to 310, we can find some slight increase.

```{r}
# change max
walk_out_money <- rep(NA, 100)
m <- matrix(nrow = 100, ncol = 32)

for(i in seq(1, 32)){
  for(j in seq_along(walk_out_money)){
    m[j, i] <- one_series(B= 200, W = 300, L = 1000, M = 10 * (i-1)) %>% get_last
  }
}

# Estimated earnings
mean_earnings <- colMeans(m) - 200

df4 <- as.data.frame(mean_earnings)
df4["Maximum_wager"] <- seq(0,310,10)
df4 %>%
  ggplot(mapping = aes(x = Maximum_wager, y = mean_earnings)) +
  geom_point() +
  geom_smooth()

# plot earnings
#plot(mean_earnings)
```


### 3. The average number of plays is at around 200.

In order to estimate the average number of plays before stopping, we need to find the number of plays in each series of play. We can easily know that the length of the one_series function represents the number of plays. This is because one_series function return the vector which stores all the budget over series of plays. If we count the number of budget, we can get the number of plays in this series.

Then, we can simulate 100 times and average the results.


```{r}
# Q4: average number of plays
# Simulation
set.seed(123)
num_play <- function(x) length(x) 

walk_out_play <- rep(NA, 100)
for(j in seq_along(walk_out_play)){
  walk_out_play[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% num_play
}

## num_play, we can easily use length. they are the same

# Walk out play distribution
hist(walk_out_play, breaks = 100)



# Estimated playing times
mean(walk_out_play)

```
### 4.1 Limitations of the simulation:

1) When we do the simulation, we need a good theory and causal hypotheses. 

2) The use of simulation requires very little time and resources. We don not need to do the experiment and collect the data in the real world. The only thing we need to do is to choose a theory and set up the parameters. But there still exists some skepticism. This is because most of today's analysis is based on reporting and building statistical models to describe what has happened in the past.

3) Even though simulation doesn't require data, the validation of simulation often requires multiple data sources to achieve the confidence in its representation of real-world dynamics.

4) Simulation need to take a lot of data to be input, so it usually need the computer with higher memory and processor speed.

### 4.2 Sources of uncertainty

1) The player's risk preference will affect the result. If he prefers risk aversion, he will stop playing afer losing a small amount of money.

2) The player's mood is changing in the whole process of playing. This will affect the strategy he choose to use.



