---
title: "06-06-order-statistic"
date: "9/29/2019"
author: "Yasi Wang"
output: html_document
---  
I haven't finished yet. I'll finish by Tuesday.

The median is an important quantity in data analysis. It represents the middle value of the data distribution. Estimates of the median, however, have a degree of uncertainty because (a) the estimates are calculated from a finite sample and (b) the data distribution of the underlying data is generally unknown. One important roles of a data scientist is to quantify and to communicate the degree of uncertainty in his or her data analysis.

Today, I'll demonstrate the empirical and theoretical methods to show the order statistics' properties.

# Density for the median
Begin with the median from a sample of N=200 from the standard normal distribution. The density function for the median in this sample is listed below. 

**Note:** the 100th order statistic is approximately the median.
```{r}
dorder <- function(x){
  100*
  choose(200,100)*
  (pnorm(x))^(100-1)*
  (1-pnorm(x))^(200-100)*
  dnorm(x)  
}

curve(
  dorder(x)
  , xlim = c(-0.5, 0.5)
  , xlab = parse(text = "X[100]")
  , ylab = "Density"
  , lwd = 3
  , col = "orange"
  , main = "Density for the Median"
)

```
From the graph, we can find that if the median increase from negative infinite to 0, the density for this median will increase from 0 to 4., the probability 

In the standard normal distribution, we can see that from the 5<sup>th</sup> to 50<sup>th</sup> quantile, the length will decrease, and from 50<sup>th</sup> to 95<sup>th</sup> quantile, the length will then increase. This means the 50% quantile has the shortest length with the most precision in the simulation. And the length gets larger with less presicion when away from 50<sup>th</sup> quantile.

Q: Write an R function that is the probability function for the median in this sample. Use the order statistic formula discussed in class. Generate a plot of the function.
```{r}
porder <- function(x){
  pbinom(100-1, 200, pnorm(x), lower.tail = FALSE)
}

curve(
  porder(x)
  , xlim = c(-0.5, 0.5)
  , xlab = parse(text = "X[100]")
  , ylab = "Probability"
  , lwd = 3
  , col = "orange"
  , main = "Probability for the Median"
)

```

Q: Write an R function that is the quantile function for the median in this sample. (You have several options for how to write this function.) Generate a plot of the function.

```{r}
qorder <- function(p) {
  out <- p
  for (i in seq_along(p)) {
    pnew <- function(x) {
      pbinom(100-1, 200, pnorm(x), lower.tail = FALSE) - p[i]
    }
    out[i] <- uniroot(pnew, c(-10,10))$root
  }
  out
}

p <- seq(0.001, 0.999, 0.001)
plot(p, qorder(p),
  , xlim = c(0, 1)
  , xlab = parse(text = "Quantile")
  , ylab = "Probability"
  , type = "l"
  , lwd = 3
  , main = "Quantile for the Median"
  , col = "orange"
)

curve(qorder(x)
  , xlab = parse(text = "Quantile")
  , ylab = "Probability"
  , type = "l"
  , lwd = 3
  , main = "Quantile for the Median"
  , col = "orange"
)
```


Q: Simulate the sampling distribution for the median as you did in the previous deliverable. Create a plot of the empirical CDF (ECDF). Overlay the plot of the ECDF with a plot of the CDF.

```{r}
set.seed(123)

sim_median <- c()
for (i in 1:5000) {
  sim_median[i] <- median(rnorm(200))
}

sim_median_ecdf <- ecdf(sim) 
plot(sim_median_ecdf, do.points = FALSE, col = "red", lwd = 3, main = "ECDF and CDF")
curve(porder(x), add = TRUE, lwd = 3, col = "blue")
legend(
    "topleft"
  , c("ECDF","CDF")
  , lwd = 3
  , col = c("red","blue")
  , bty = "n"
)

```


Q: Using the simulated sampling distribution from the previous question, create a histogram (on the density scale). Overlay the histogram with a plot of the density function.

```{r}
hist(sim_median, breaks = 30, col = "lightgray", freq = FALSE)
lines(density(sim_median), col = "orange", lwd = 3)
```


Q: One very common way to compare a random sample to a theoretical candidate distribution is the QQ plot. It is created by ploting quantiles of the theoretical distribution on the x-axis and empirical quantiles from the sample on the y-axis.

```{r include=FALSE}
# library(devtools)
# install_github("thomasgstewart/tgsify")
library(tgsify)
```


If sample and theoretical quantiles come from the same distribution, then the plotted points will fall along the line y=x, approximately. 


I generated a QQ plot for the simulated data of the median relative to the known sampling distribution of the median.
Does the simulated data agree with the theoretical sampling distribution?
```{r}
x1 <- qorder((1:199)/200)
y1 <- quantile(sim_median, probs = (1:199)/200)

plotstyle(style = upright)
plot(x1,y1, asp = 1, xlab = "Theoretical quantile", ylab = "Sample quantile")
abline(0,1)
```
Clearly, they are from the same distribution



Q: Modify the dorder, porder, and qorder functions so that the functions take a new parameter k (for the kt**h order statistic) so that the functions will work for any order statistic and not just the median.
```{r}
dorder_new <- function(k, x){
  k*
  choose(200,k)*
  (pnorm(x))^(k-1)*
  (1-pnorm(x))^(200-k)*
  dnorm(x)  
}

porder_new <- function(k, x){
  pbinom(k-1, 200, pnorm(x), lower.tail = FALSE)
}

qorder_new <- function(k, p) {
  out <- p
  for (i in seq_along(p)) {
    pnew <- function(x) {
      pbinom(k-1, 200, pnorm(x), lower.tail = FALSE) - p[i]
    }
    out[i] <- uniroot(pnew, c(-10,10))$root
  }
  out
}
```


Q: Generate the QQ plot for simulated data from the sampling distribution of the sample max and the theoretical largest order statistic distribution.

```{r}
set.seed(123)

sim_max <- c()
for (i in 1:5000) {
  sim_max[i] <- max(rnorm(200))
}

x2 <- qorder_new(k = 200, (1:199)/200)
y2 <- quantile(sim_max, probs = (1:199)/200)

plotstyle(style = upright)
plot(x2,y2, asp = 1, xlab = "Theoretical quantile", ylab = "Sample quantile")
abline(0,1)
```

Q: Modify the dorder, porder, and qorder functions so that the functions take new parameters dist and ... so that the functions will work for any continuous distribution that has d and p functions defined in R.
```{r}

require(dplyr)
require(data.table)
`%|%` <- function(a,b) paste0(a,b)
require(tgsify) #can be installed with devtools::install_github("thomasgstewart/tgsify")


dist = c("norm", "exp", "binom", "unif", "beta")

dorder <- function(dist, k, x, n){
  dist = c("norm", "exp", "binom", "unif", "beta")
  pf <- get("p" %|% dist)
  df <- get("d" %|% dist)
  
  k*
  choose(n,k)*
  (pf(x))^(k-1)*
  (1-pf(x))^(n-k)*
  df(x)  
}

porder <- function(k, x, n){
  pbinom(k-1, n, pf(x), lower.tail = FALSE)
}

qorder <- function(k, x, n) {
  out <- x
  for (i in seq_along(x)) {
    pnew <- function(x) {
      pbinom(k-1, n, pf(x), lower.tail = FALSE) - p[i]
    }
    out[i] <- uniroot(pnew, c(-100,100))$root
  }
  out
}



pf <- pnorm
df <- dnorm


dist = c("norm", "exp", "binom", "unif", "beta")

rorder <- get("r" %|% dist)


dorder <- function(dist)

dorder <- function(k, x, n){
  k*
  choose(n,k)*
  (pnorm(x))^(k-1)*
  (1-pnorm(x))^(n-k)*
  dnorm(x)  
}

one_setting <- function(params, ps = seq(.05, .95, by = 0.05)){
  inputs <- as.list(params)
  inputs$ps <- ps
  with(inputs, {
    N_p <- ps %>% length
    pf <- get("p" %|% dist)
    rf <- get("r" %|% dist)
    qf <- get("q" %|% dist)
    df <- get("d" %|% dist)

  })
}

    
    

sim_settings <- expand.grid(
    N = c(200, 400, 800, 1600)
  , M = 5000
  , dist = c("norm", "exp", "binom", "unif", "beta")
  , KEEP.OUT.ATTRS = FALSE
  , stringsAsFactors = FALSE
)

results <- vector("list", nrow(sim_settings))


one_setting <- function(params, ps = seq(.05, .95, by = 0.05)){
  inputs <- as.list(params)
  inputs$ps <- ps
  with(inputs, {
    N_p <- ps %>% length
    rf <- get("r" %|% dist)
    qf <- get("q" %|% dist)
    df <- get("d" %|% dist)
    
    # Array option
    # vals <- array(rf(N*M), dim = c(M,N))
    # q1 <- vals %>% apply(1, quantile, probs = ps)
    
    # For loop option
    q1 <- array(NA, dim = c(N_p, M))
    for(i in 1:M){
      q1[,i] <- rf(N) %>% quantile(probs = ps)
    }
    
    q2 <- q1 %>% apply(1, quantile, probs = c(0.025, 0.975), type = 4)
    ls <- q2 %>% apply(2, diff)
    
    # p to q to d
    density <- ps %>% qf %>% df
    
    #output
    data.frame(p = ps, density = density, length = ls, N = N, M = M, dist = dist, stringsAsFactors = FALSE)
  })
}

for(i in 1:nrow(sim_settings)){
  results[[i]] <- sim_settings[i,] %>% one_setting  
}

ar <- rbindlist(results) %>% 
  mutate(N_color = N %>% as.factor %>% as.numeric) 


```


Q: Use the newly modified functions to plot the probability and density functions for the sample min (N=200).

```{r}

```

