---
title: "Final Exam"
output: html_document
---

0. Names: Yasi Wang

# 1. Simulation

The Monte Hall problem is a classic game show.  Contestants on the show where shown three doors.  Behind one randomly selected door was a sportscar; behind the other doors were goats.

At the start of the game, contestants would select a door, say door A.  Then, the host would open either door B or C to reveal a goat.  At that point in the game, the host would ask the contestant if she would like to change her door selection.  Once a contestant decided to stay or change, the host would open the choosen door to reveal the game prize, either a goat or a car.

In this problem, consider a **modified** version of the Monte Hall problem in which the number of doors is **variable**.  Rather than 3 doors, consider a game with 4 or 5 or 50 doors.  In the modified version of the game, a contestant would select an initial door, say door A.  Then, the host would open **one** of the remaining doors to reveal a goat.  At that point in the game, the host would ask the contestant if she would like to change her door selection.  Once a contestant decided to stay or change, the host would open the choosen door to reveal the game prize, either a goat or a car.

Consider two strategies:
  
  1. Always stay with the first door selected.
  2. Always switch to the unopened door.

**C.** The function `game` below plays a single game of Monte Hall.  The function returns a vector of length two, the first element is the prize under strategy 1 and the second element is the prize under strategy 2.  The function has a single input parameter, N, which is the number of doors in the game.

Use the `game` function to estimate the probability that both strategies result in a goat. Let **N=4**.

```{r}
require(magrittr)
require(dplyr)

game <- function(N){
  if(N<3) stop("Must have at least 3 doors")
  prize <- sample(c(rep("goat",N-1),"car"), N)
  guess <- sample(1:N,1)
  game <- data.frame(door = 1:N, prize = prize, stringsAsFactors = FALSE) %>% 
    mutate(first_guess = case_when(
      door == guess ~ 1
      , TRUE ~ 0
    )) %>% 
    mutate(potential_reveal = case_when(
        first_guess == 1 ~ 0
      , prize == "car" ~ 0
      , TRUE ~ 1
    )) %>% 
    mutate(reveal = 1*(rank(potential_reveal, ties.method = "random") == 3)) %>% 
    mutate(potential_switch = case_when(
      first_guess == 1 ~ 0
      , reveal == 1 ~ 0
      , TRUE ~ 1
    )) %>% 
    mutate(switch = 1*(rank(potential_switch, ties.method = "random") == 3))
  c(game$prize[game$first_guess == 1], game$prize[game$switch == 1])
}
```


```{r}
cnt1 = 0
cnt2 = 0

out <- matrix(NA, nrow = 5000, ncol = 2)

for (i in 1:5000) {
  out[i,] = game(4)
  if (out[i,1] == "goat" & out[i, 2] == "goat") {
    cnt1 = cnt1 + 1
  } else {
    cnt2 = cnt2 + 1
  }
}

p1 <- cnt1/5000
p1

```

`answer:` the probability that both strategies result in a goat is p1 =  0.3754.


**B**. Continuing from part **C**, what is the probability that at least one of the strategies results in winning a car?

```{r}
p2 <- 1 - p1
p2
```
`answer:` the probability that at least one of the strategies results in winning a car is p2 = 0.6246.

**A**. Communicate the precision of your simulated probability in part **B** by calculating a **99\%** confidence interval.

```{r}

sum(out[,1] == "car"|out[,2]=="car")

prop.test(3138, 5000, 5/8, conf.level = 0.99) # the true prop doesn't matter. we only want the 99% CI

```

`answer:` the 99% CI is c( 0.610, 0.645).

# 2. Probability

Consider a test for a rare genetic condition.  Let T+ denote a test result that indicates the condition is present, while T- denotes absence.  Let D+ and D- denote the true status of the disease.

**C**.  Fill-in the probability table using the following information:

+ P(T+|D+) = .85,  and 
+ P(T-|D-) = .95,  and 
+ P(D+) = 0.001

|    |    D+   |    D-  |        |
|:--:|:-------:|:------:|:------:|
| T+ | 0.00085 | 0.04995| 0.0508 |
| T- | 0.00015 | 0.94905| 0.9492 |
|    |  0.001  |  0.999 |    1   |



**B**. Calculate the **negative** predictive value of the test, P(D-|T-).

```{r}
p_negative <- 0.94905/0.9492
p_negative

```
`answer:` P(D-|T-) = P(D- and T-)/P(T-) = 0.999842

**A** Create a plot that shows how the **positive** predictive value as a function of the prevalence of disease, P(D+).


|    |   D+  |    D-    |          |
|:--:|:-----:|:--------:|:--------:|
| T+ | 0.85x | 0.05(1-x)| 0.8x+0.05|
| T- | 0.15x | 0.95(1-x)| 0.95-9.8x|
|    |    x  |    1-x   |      1   |

P(D+|T+) = P(D and T)/P(T)
```{r}
prevalence <- seq(0.001, 0.1, length = 50)
ppv <- (0.85*prevalence)/(0.8*prevalence+0.05)

plot(prevalence, ppv, xlab = "Prevalence", ylab = "PPV")
```

# 3. Discrete Distributions

Suppose the yearly hospital charges (in thousands of dollars) for a randomly selected Vanderbilt student is a mixture distribution.

For 50% of students, the hospital charges will be $0.  For the remaining 50% of students, the hospital charges are a random variable described by a gamma distribution with shape = 2 and scale = 2.  (Again, in thousands of dollars.)   

```{r}
hospital_charges <- function(N){
  group <- rbinom(N, 1, 0.5)
  charges <- 0*group + rgamma(N, shape = 2, scale = 2)*(1-group)
  charges
}
```

**C**.  What is the 90th percentile for yearly hospital charges for a randomly selected Vanderbilt student?

```{r}
out <- NA
for (i in (1:5000)){
  trial <- hospital_charges(5000)
  out[i] <- quantile(trial, 0.9)
}

mean(out)

```

`answer:` the 90th percentile is 5.987764 ((in thousands of dollars).

**B**.  Consider the **class** average yearly hospital charge for the students in a class of size 30.  Plot the density function or a simulated histogram of the class average yearly hospital charge.

```{r}
out <- NA
for (i in (1:5000)){
  trial <- hospital_charges(30)
  out[i] <- mean(trial)
}

hist(out, main = "Histogram of the class average yearly hospital charge", xlab = "Thousands of dollars")

```


**A**.  What is the probability that a randomly selected class of size 30 students will have less than 10 students with zero yearly hospital charges?


```{r}
total = 0

for (j in (1:5000)){
  cnt = 0
  for (i in (1:30)){
    res <- hospital_charges(1)
    if (res == 0){
    cnt = cnt + 1
    }
  }
  if (cnt < 10) {
    total = total + 1
  }
}

total/5000


```
`answer:`the probability is 0.0206.

# 4. Continuous Distributions

**C.** Suppose diastolic blood pressure (DBP) follows a normal distribution with mean 80 mmHg and SD 15 mmHg. What is the probability that a randomly sampled person's DBP lies between 70 and 104 mmHg?

```{r}
pnorm(104, mean = 80, sd = 15) - pnorm(70, mean = 80, sd = 15)

```

`answer:` p = 0.6927

**B.** Suppose a human femur was discovered that is 37 cm long.  Also suppose that using the NHANES data, researchers believe the distribution of femor bones, by sex, are distributed as follows:

+ Female adult femor $\sim N(36, 3.3)$
+ Male adult femor $\sim N(40, 3.4)$


Under the assumption that male and females are equally likely, what is the probability that the discovered femor was from a male?

```{r}

# bayes rule: P(A|B) = P(B|A)P(A)/(P(B|A)P(A) + P(B|~A)P(~A))

dnorm(37,40,3.4)*0.5 /((dnorm(37,40,3.4)*0.5) + (dnorm(37,36,3.3)*0.5))

# 0.407765

```

`answer:` p = 0.408.

**A.**  Continuing part **B**, generate a plot of P(femor from male | femor length = x).  Let femor length range from 25 to 50.

```{r}
femor_length <- 25:50
prob_male <- dnorm(femor_length,40,3.4)*0.5 /((dnorm(femor_length,40,3.4)*0.5) + (dnorm(femor_length,36,3.3)*0.5))

plot.new()
plot.window(xlim = c(25,50), ylim = c(0,1))
lines(femor_length, prob_male)
axis(1)
axis(2)
box()
title(xlab = "Femor Length", ylab = "P( Male | femor length)")
```

# 5. Expectation and Variance

Let us revisit the yearly hospital charges distribution from a previous section.

>**Recall:** The yearly hospital charges (in thousands of dollars) for a randomly selected Vanderbilt student is a mixture distribution. For 50% of students, the hospital charges will be $0.  For the remaining 50% of students, the hospital charges are a random variable described by a gamma distribution with shape = 2 and scale = 2.  (Again, in thousands of dollars.)   

```{r}
hospital_charges <- function(N){
  group <- rbinom(N, 1, 0.5)
  charges <- 0*group + rgamma(N, shape = 2, scale = 2)*(1-group)
  charges
}
```

**C.** What is E[yearly hospital charges]?

```{r}
mean(hospital_charges(10000))
```

`answer:` E[yearly hospital charges] = 2.03 (thousands of dollors).

**B.** Suppose Vanderbilt implements a cap of \$10,000 on yearly student hospital charages.  What is the mean yearly hospital charge under the new policy?

```{r}

out <- 0

for (i in (1:100000)){
  one_time <- hospital_charges(1)
  if (one_time > 10){
    one_time = 10
  }
  out <- out + one_time
}

mean(out)/100000


```

`answer:` E[yearly hospital charges in new policy] = 1.96 (thousands of dollors).

**A.** What is the variance of yearly hospital charge under the new policy?
```{r}

# method1: Var(X)=E((X???E(X))^2)
var <- 0
for (i in (1:100000)){
  one_time <- hospital_charges(1)
  if (one_time <= 10) {
  var <- var + (one_time - 1.96)^2
  } else if (one_time > 10){
    var <- var + (10 - 1.96)^2
  }
}
var/100000


# method2: var(x) = E(x^2) - E(x)^2


```
`anwer:` the variance of yearly charge in new policy is 6.96 (thousands of dollars)

# 6. Transformations & Sampling Distributions

**C.** Consider the log normal distribution.  If X is a log normal random variable, then log(X) is a normal random variable.  One way to create pseudo-random draws from the log normal distribution is to generate draws from a normal distribution and then to transform the draws by expononentiating.  The parameters of the log normal distribution are the parameters of the underlying normal distribution, $\mu$ and $\sigma$ (or $\sigma^2$).  

Log normal data are prevalent is biological systems and econometrics.

Suppose a blood chemistry measure has a log normal distribution with $\mu$ = 0 and $\sigma$ = 1. Generate an histogram or density curve for the sampling distribution of the median when the sample size is 101.

```{r}
out <- NA
for (i in 1:5000){
  out[i] <- rlnorm(101, 0, 1) %>% median
}

hist(out, prob = T, main= "Histogram/Density of the sampling distribution of the median", ylim = c(0, 3.5))
lines(density(out), col = "red", lwd = 3)

```


**B.** Below is the CDF function for the kth order statistic when the underlying distribution is log normal with $\mu$ = 0 and $\sigma$ = 1.  Create a plot of the ECDF of the simulated sampling distribution generated in **C** and overlay the CDF using the function below.

```{r}
Fk <- function(x,k,n){
  pbinom(k-1, n, plnorm(x), lower.tail = FALSE)
}

m <- ecdf(out)

plot(m, do.points = FALSE, col = "red", lwd = 3)
curve(Fk(x, 51, 101), add =T, col = "blue", lwd = 3)
legend(
    "topleft"
  , c("ECDF","CDF")
  , lwd = 3
  , col = c("red","blue")
  , bty = "n"
)

```

**A.** Of the 25th, 50th, and 75th quantiles of the distribution from **B**, which will have the tighest 95% CI?  (Show the sampling distribution of each.)

```{r}

# Simulation method
out1 <- NA
out2 <- NA
out3 <- NA
for (i in 1:5000){
  sample <- rlnorm(101, 0, 1)
  out1[i] <- quantile(sample, 0.25)
  out2[i] <- quantile(sample, 0.5)
  out3[i] <- quantile(sample, 0.75)

}

quantile(out1, c(0.05/2, 1-0.05/2))
quantile(out2, c(0.05/2, 1-0.05/2))
quantile(out3, c(0.05/2, 1-0.05/2))

plot(density(out1), col = "red", xlim = c(0.3, 3.5), lwd = 3, main = "Distribution of 25th, 50th, 75th quantiles")
lines(density(out2), col = "blue", lwd = 3)
lines(density(out3), col = "orange", lwd = 3)
legend(
    "topright"
  , c("25th","50th", "75th")
  , lwd = 3
  , col = c("red","blue", "orange")
  , bty = "n"
)

```
`answer:` According to the plot and results for CI, we can find that 25th quantile has the tighest 95% CI with [0.392, 0.670]


# 7. Estimation of CDF and PDF from data

The following code will load the NHANES data and select the first 500 rows.

```{r}
Hmisc::getHdata(nhgh)
d1 <- nhgh[1:500,]
```

**C.** Estimate the distribution of standing height for adult (age > 18) males using the MLE method with a normal distribution.  Create a plot of the estimated density function.

```{r}

# in normal distribution, MLE is the mean and sd of the variable
ht_male <- d1 %>% 
  filter(age > 18) %>% 
  filter(sex == "male") %>% 
  pull(ht)
mean_mle <- mean(ht_male)
sd_mle <- sd(ht_male)


hist(ht_male, freq = FALSE, ylim = c(0, 0.06), main = "Estimated standing height for adult males using MLE", xlab = "height")
curve(dnorm(x, mean_mle, sd_mle), add = TRUE, lwd = 3, col = "red")

```


**B.** Estimate the distribution of BMI for adult (age > 18) females using using the method of moment method with the gamma distribution. Create a plot of the estimated density function.

```{r}
bmi_female <- d1 %>% 
  filter(age > 18) %>% 
  filter(sex == "female") %>% 
  pull(bmi)

xbar <- mean(bmi_female)
s2 <- var(bmi_female)

shape_hat <- xbar^2/s2 
scale_hat <- s2/xbar


curve(dgamma(x, shape = shape_hat, scale = scale_hat), xlim = c(0, 100), col = "red", lwd = 3, main = "Estimated BMI for adult female using Method of Moment method", ylab = "Density", xlab = "Female BMI")

```


**A.** Estimate the distribution of creatinine (SCr) for adults (age > 18) using the kernel density method with a gaussian kernel.  Create a plot of the estimated density function.

```{r}
scr <- d1 %>% 
  filter(age > 18) %>% 
  filter(!is.na(SCr)) %>% 
  pull(SCr)

# f^     density function
epdfstar <- function(t, data, smooth){
  outer(t, data, function(a,b){ dnorm(a, b, smooth)}) %>% rowMeans
}

hist(scr, main = "EPDF using kernal density estimate", freq = FALSE, xlim = c(-2, 8), ylim = c(0, 1))
curve(epdfstar(x, scr, smooth = 0.5), add = TRUE, lwd = 3, col = "blue")
# second way: defaul density function

# den1 <- density(scr, adjust = 1.5, kernel = "gaussian")
# lines(den1, lwd = 3, col = "red")


```

# 8. Sample from an estimated distribution

The following code will load the low birth weight data from the MASS package.  The description of the variables in the dataset can be found in the birthwt documentation with the command `?MASS::birthwt`.

```{r}
bwt <- MASS::birthwt
```

**C.** Generate a 95% confidence interval for the mean birthweight of infants whose mothers **did** smoke during pregnancy using the bootstrap.
```{r}

sm <- bwt %>% 
  filter(smoke == 1) %>% 
  filter(!is.na(bwt)) %>% 
  pull(bwt)
# bootstrap

R <- 5000
out <- rep(NA, R)
for (i in 1:R){
  out[i] <- sample(sm, length(sm), replace = T) %>% mean
}

quantile(out, c(0.05/2, 1-0.05/2))

```

`answer:` the 95% CI is [2625.484, 2916.043]

**B.** Generate a 95% confidence interval for the mean birthweight of infants whose mothers **did** smoke during pregnancy using the Central Limit Theorem shortcut.

```{r}

# t.test
t.test(sm, data = bwt)

# by hand

mean_bwt <- mean(sm) 
std_bwt <- sd(sm)
N <- length(sm)

Z1 <- -1.96
Z2 <- 1.96

ci_left <- Z1 * (std_bwt/sqrt(N)) + mean_bwt
ci_right <- Z2 * (std_bwt/sqrt(N)) + mean_bwt

ci_left
ci_right

```

`answer:` the 95% CI is [2621.624, 2922.214]

**A.** Let $\mu_s$ be the mean birthweight of infants whose mothers smoked during pregnancy.  Let $\mu_{ns}$ be the mean for the non-smoking group.  Use simulation to calculate the 95% confidence interval for $\mu_s/\mu_{ns}$.

```{r}

nonsm <- bwt %>% 
  filter(smoke == 0) %>% 
  filter(!is.na(bwt)) %>% 
  pull(bwt)
  

out <- NA
for (i in 1:1000){
  out1 <- sample(sm, length(sm), replace = T) %>% mean
  out2 <- sample(nonsm, length(nonsm), replace = T) %>% mean

  out[i] <- out1/out2
}

quantile(out, c(0.05/2, 1-0.05/2))


```

`answer:` the 95% CI for $\mu_s/\mu_{ns}$ is [0.846, 0.973].

# 9.  Inference

**C.** Suppose two studies were performed looking at the risk of mild complication after hernia repair using open and laparoscopic surgical approaches.  The study results are below.  Using the data from each study individually, perform the hypothesis test that the risk of complication between open and laparoscopic repairs are the same under the usual point null. What is the p-value from each study?  What do you conclude from each study?


| Study 1 | Comp | No comp |
|:---|:---|:---|
| Open | 30 | 70 |
| Lap  | 35 | 65 |

| Study 2 | Comp | No comp |
|:---|:---|:---|
| Open | 600 |     1400 |
| Lap  | 619 |     1381 |

```{r}
prop.test(c(30, 35), c(30+70, 35+65))

prop.test(c(600, 619), c(600+1400, 619+1381))

```
`answer:` In the first study, p = 0.5459. In the second study, p = 0.5364. We can find that in both study, 0 is included in the 95% CI. We conclude that in both study, we fail to reject the null under alpha = 0.05 level, which means that the risk of complication between open and laparoscopic repairs are the same.


**B.** Suppose that prior to the studies, the researchers established an equivalence threshold of 6 percentage points.  Using the confidence intervals, which studies (if any) showed a conclusive similarity between surgical approaches for the complication rate.  Explain why.

`answer:` The second study showed a conclusive similarity between surgical approaches for the complication rate. Because we can find that 95% CI in first study is [-0.190, 0.090] while in the second study is [-0.039,  0.020]. The lengths in the two invervals are 0.280 and 0.059, respectively. Thus, the second inverval is inside of the 6 percentage points, which means that it shows a conclusive similarity in this way.


**A.** If the data from the studies were combined, what is the smallest equivalence threshold that would identify a conclusive similarity between the surgical approaches?

```{r}
res <- prop.test(c(630, 654), c(630+1470, 654+1446))

threshold <- res$conf.int[2] - res$conf.int[1]
threshold

```

`answer:` The smallest equivalence threshold should be 5.67 percentage points

# 10.  Joint Distributions

**C.** Fill in the blank.  The sample correlation is a measure of ____`linear`___ association.

**B.** Explain why predictions from a conditional distribution generally have smaller prediction error than predictions from the marginal distribution.

`answer:` If two variable are correlated, conditional distribution has less variance than marginal, so we can predict it with less error. If two variable are not correlated, then conditional distribution has no advantage than marginal in prediction.

Let's look at MSE (MSE = VAR + BIAS^2). If the marginal model and the conditional model both have zero bias, then the difference in MSE comes does to variance. The variance of the conditional model is less than or equal to the variance of the marginal model.  This means that the MSE of the conditional model will be less than or equal to the MSE of the marginal model.


**A.** Use the CLT shortcut to calculate the 95% confidence interval for the correlation of arm circumferance and arm length using the NHANES dataset.  Is the sample correlation a reasonable measure of association for this data?

```{r}
#short cut
Hmisc::getHdata(nhgh)

df <- nhgh

cor.test(df$armc, df$arml)

```
`answer:` Our null hypothesis is that the arm circumferance and arm length has no correlation. Using the cor.test (CLT short cut), we can find that the sample estimated correlation between arm circumferance and arm length is 0.450, which falls in the 95% CI (0.48, 0.52). The p-value is smaller than 0.05. We can conclude that, in the alpha = 0.05 level, we fail to reject the null, which  means that the sample correlationa is a reasonable measure of association for this data.

